---
title: "Data Analysis with ctSEM models"
author: "Clara Behnke"
format: 
  pdf:
    code-block-border-left: false
execute: 
  warning: false
---

## Load packages and data set

```{r}
library(readr)
library(tidyverse)
library(ctsem)
```

```{r}
dat_aggregated <- read_csv("aggregated_data_triggered_ZI.csv")
```

## Grand mean centering and scaling

```{r}
dat_aggregated_c <- dat_aggregated %>%
  mutate(across(c("A1a", "A1b", "A3", "A9", "A4", "K3", "K4", "A8", "A6", "A2", "A5", "A7"), scale)) 
```

## Single participant

```{r}
# participant with highest number of completed questionnaires
ID_fewestNAs <- dat_aggregated %>%
  filter(interactionStatus == "Completed") %>%
  count(pseudonym, sort = TRUE) %>%
  slice(1) %>%
  pull(pseudonym)

dat_fewestesNAs <- dat_aggregated %>%
  filter(pseudonym == ID_fewestNAs)

# individual centering and scaling
dat_fewestesNAs_c <- dat_fewestesNAs %>%
  mutate(across(c("A1a", "A1b", "A3", "A9", "A4", "K3", "K4", "A8", "A6", "A2", "A5", "A7"), scale)) 

# Model specification
model_1 <- ctModel(type = "stanct", time = "time_hours", id = "pseudonym",
                   n.latent = 3, latentNames = c("symptom_load", "acceptance", "defusion"),
                   n.manifest = 3, manifestNames = c("A1b", "A2", "A3"),
                   LAMBDA=diag(3))
model_1$pars$indvarying <- FALSE

# Model fitting
fit_1 <- ctStanFit(datalong = dat_fewestesNAs_c, ctstanmodel = model_1, priors = FALSE, optimize = TRUE) 
# priors: True -> HMC, false -> max likelihood

# Summary
summary(fit_1)
summary(fit_1)$popmeans

# continuous parameters
ctStanContinuousPars(fit_1, calcfunc = quantile,
  calcfuncargs = list(probs = 0.5))$DRIFT
ctStanContinuousPars(fit_1, calcfunc = mean)$DRIFT

plot(fit_1, wait = FALSE)

ctStanDiscretePars(fit_1, plot = TRUE, # regression coefficient for particular time intervals
                   indices = 'CR') # cross effects (AR would be auto effects)

#ctStanPlotPost(obj = fit_1, rows = 3) # posterior vs prior (here no priors)

#ctKalman(fit_1,
#         timerange = "asdata",
#         timestep = "auto",
#         plot = TRUE, 
#         plotcontrol = list(xaxs = 'i', main = "Predicted"))

```


## All participants and ACT modules, no priors

```{r}
model_all <- ctModel(type = "stanct", time = "time_hours", id = "pseudonym",
                     n.latent = 7, 
                     latentNames = c("symptom_load", "acceptance", "defusion", "presence", "self_context", "values", "commited_action"),
                     n.manifest = 7, 
                     manifestNames = c("A1b", "A2", "A3", "A4", "A5", "A6", "A7"),
                     LAMBDA = diag(7))

# Model fitting
fit_all <- ctStanFit(datalong = dat_aggregated_c, ctstanmodel = model_all, priors = FALSE, optimize = TRUE) 

# Summary
summary(fit_all)

summary(fit_all)$popmeans

# continuous parameters
ctStanContinuousPars(fit_all, calcfunc = quantile,
  calcfuncargs = list(probs = 0.5))$DRIFT
ctStanContinuousPars(fit_all, calcfunc = mean)$DRIFT

ctStanDiscretePars(fit_all, plot = TRUE) + theme(legend.position = "none")
ctKalman(your_model, plot = TRUE)

ctStanDiscretePars(fit_all, plot = TRUE, # regression coefficient for particular time intervals
                   indices = 'CR') # cross effects (AR would be auto effects)

```

## All participants and ACT modules, with priors

```{r}
#| cache: true

model_all <- ctModel(type = "stanct", time = "time_hours", id = "pseudonym",
                     n.latent = 7, 
                     latentNames = c("symptom_load", "acceptance", "defusion", "presence", "self_context", "values", "committed_action"),
                     n.manifest = 7, 
                     manifestNames = c("A1b", "A2", "A3", "A4", "A5", "A6", "A7"),
                     LAMBDA = diag(7))

# Model fitting
fit_all_hmc <- ctStanFit(datalong = dat_aggregated_c, ctstanmodel = model_all, priors = TRUE, optimize = FALSE) 
# priors: True -> HMC, false -> max likelihood

# Summary
summary(fit_all_hmc)

summary(fit_all_hmc)$popmeans

# continuous parameters
ctStanContinuousPars(fit_all_hmc, calcfunc = quantile,
  calcfuncargs = list(probs = 0.5))$DRIFT
ctStanContinuousPars(fit_all_hmc, calcfunc = mean)$DRIFT

plot(fit_all_hmc, wait = FALSE)

ctStanDiscretePars(fit_all_hmc, plot = TRUE, # regression coefficient for particular time intervals
                   indices = 'CR') # cross effects (AR would be auto effects)

ctStanPlotPost(obj = fit_all_hmc, rows = 3) # posterior vs prior (here no priors)

ctKalman(fit_all_hmc,
         timerange = "asdata",
         timestep = "auto",
         plot = TRUE, 
         plotcontrol = list(xaxs = 'i', main = "Predicted"))

```


## All participants: defusion -> symptom load

### Models

```{r}
# without random intercept, time scale = hours
model_defus <- ctModel(type = "stanct", time = "time_hours", id = "pseudonym",
                       n.latent = 2, 
                       latentNames = c("symptom_load", "defusion"),
                       n.manifest = 2, 
                       manifestNames = c("A1b", "A3"),
                       LAMBDA = diag(2))
model_defus$pars$indvarying <- FALSE
head(model_defus)

# without ri, time scale = days
model_defus_d <- ctModel(type = "stanct", time = "time_days", id = "pseudonym",
                       n.latent = 2, 
                       latentNames = c("symptom_load", "defusion"),
                       n.manifest = 2, 
                       manifestNames = c("A1b", "A3"),
                       LAMBDA = diag(2))
model_defus_d$pars$indvarying <- FALSE
head(model_defus_d)

# with random intercept
model_defus_ri <- ctModel(type = "stanct", time = "time_hours", id = "pseudonym",
                          n.latent = 2, 
                          latentNames = c("symptom_load", "defusion"),
                          n.manifest = 2, 
                          manifestNames = c("A1b", "A3"),
                          LAMBDA = diag(2))
head(model_defus_ri)

# with random intercept, time scale = days
model_defus_d_ri <- ctModel(type = "stanct", time = "time_days", id = "pseudonym",
                          n.latent = 2, 
                          latentNames = c("symptom_load", "defusion"),
                          n.manifest = 2, 
                          manifestNames = c("A1b", "A3"),
                          LAMBDA = diag(2))
head(model_defus_d_ri)
```

### No Random Intercept

#### ML optimization

```{r}
# Model fitting
fit_defus_ml_d <- ctStanFit(datalong = dat_aggregated_c, 
                          ctstanmodel = model_defus_d, 
                          priors = FALSE, optimize = TRUE) 

# Summary
summary(fit_defus_ml_d)

plot(fit_defus_ml_d, wait = FALSE)
```

Error in sgd(init, fitfunc = target, parsets = parsets, nsubsets = 1,  : 
  NA in parameter proposal!
  
Possible time step problem? Intervals too large? Try reduce maxtimestep

Error in (function (expr, envir = parent.frame(), enclos = if (is.list(envir) ||  : 
  Exception: quad_form_sym: A is not symmetric. A[1,2] = nan, but A[2,1] = nan  (in 'model_ctsm' at line 895)


#### MAP

```{r}
# Model fitting
fit_defus_map <- ctStanFit(datalong = dat_aggregated_c, ctstanmodel = model_defus, priors = TRUE, optimize = TRUE) 

# Summary
summary(fit_defus_map)

plot(fit_defus_map, wait = FALSE)
```

Error in (function (expr, envir = parent.frame(), enclos = if (is.list(envir) ||  : 
  Exception: mdivide_left_spd: Matrix A is not positive definite  (in 'model_ctsm' at line 924)s

```{r}
# Model fitting
fit_defus_map_d <- ctStanFit(datalong = dat_aggregated_c, ctstanmodel = model_defus_d, priors = TRUE, optimize = TRUE) 

# Summary
summary(fit_defus_map_d)

plot(fit_defus_map_d, wait = FALSE)
```

#### HMC

```{r}
#| cache: true
# Model fitting
fit_defus_hmc <- ctStanFit(datalong = dat_aggregated_c, ctstanmodel = model_defus, priors = TRUE, optimize = FALSE) 

# Summary
summary(fit_defus_hmc)

plot(fit_defus_hmc, wait = FALSE)
```

```{r}
# Diagnostics chains

plot(fit_defus_hmc, types = "trace", wait = FALSE)

rstan::stan_trace(fit_defus_hmc$stanfit$stanfit, pars = ctStanParnames(fit_defus_hmc, "pop_DRIFT"))

max(rstan::summary(fit_defus_hmc$stanfit$stanfit)$summary[, "Rhat"], na.rm = TRUE)
min(rstan::summary(fit_defus_hmc$stanfit$stanfit)$summary[, "n_eff"], na.rm = TRUE)
rstan::check_hmc_diagnostics(fit_defus_hmc$stanfit$stanfit)
```

### Random intercept

#### ML optimization

```{r}
# Model fitting
fit_defus_ml_d_ri <- ctStanFit(datalong = dat_aggregated_c, 
                               ctstanmodel = model_defus_d_ri, 
                               priors = FALSE, optimize = TRUE) 

# Summary
summary(fit_defus_ml_d_ri)

plot(fit_defus_ml_d_ri, wait = FALSE)
```

Error in delta[abs(delta) > maxparchange] <- maxparchange * sign(delta[abs(delta) >  : 
  NAs are not allowed in subscripted assignments
  
Possible time step problem? Intervals too large? Try reduce maxtimestep

Error in (function (expr, envir = parent.frame(), enclos = if (is.list(envir) ||  : 
  Exception: quad_form_sym: A is not symmetric. A[1,2] = nan, but A[2,1] = nan  (in 'model_ctsm' at line 779)



#### MAP

```{r}
#| cache: true
# Model fitting
fit_defus_map_ri <- ctStanFit(datalong = dat_aggregated_c, 
                              ctstanmodel = model_defus_ri, 
                              priors = TRUE, optimize = TRUE) 

# Summary
summary(fit_defus_map_ri)

plot(fit_defus_map_ri, wait = FALSE)
```

#### HMC

```{r}
#| cache: true
# Model fitting
fit_defus_hmc_ri <- ctStanFit(datalong = dat_aggregated_c, 
                              ctstanmodel = model_defus_ri, 
                              priors = TRUE, optimize = FALSE) 

# Summary
summary(fit_defus_hmc_ri)

plot(fit_defus_hmc_ri)
```

```{r}
# Diagnostics chains

plot(fit_defus_hmc_ri, types = "trace", wait = FALSE)

rstan::stan_trace(fit_defus_hmc_ri$stanfit$stanfit, pars = ctStanParnames(fit_defus_hmc, "pop_DRIFT"))

max(rstan::summary(fit_defus_hmc_ri$stanfit$stanfit)$summary[, "Rhat"], na.rm = TRUE)
min(rstan::summary(fit_defus_hmc_ri$stanfit$stanfit)$summary[, "n_eff"], na.rm = TRUE)
rstan::check_hmc_diagnostics(fit_defus_hmc_ri$stanfit$stanfit)
```

## Model comparison

### ML

```{r}
# without ri, time scale = days
model_defus_d_0 <- ctModel(type = "stanct", time = "time_days", id = "pseudonym",
                           n.latent = 2, 
                           latentNames = c("symptom_load", "defusion"),
                           n.manifest = 2, 
                           manifestNames = c("A1b", "A3"),
                           LAMBDA = diag(2),
                           DRIFT = matrix(0, nrow = 2, ncol = 2))
head(model_defus_d_0)

# Model fitting
fit_defus_ml_ri_0 <- ctStanFit(datalong = dat_aggregated_c, 
                              ctstanmodel = model_defus_d_0, 
                              priors = FALSE, optimize = TRUE) 

# Summary
summary(fit_defus_ml_ri_0)

# model comparison
ctChisqTest(fit_defus_ml_d_ri, fit_defus_ml_ri_0)

```

### MAP

```{r}
# without ri, time scale = days
model_defus_d_0 <- ctModel(type = "stanct", time = "time_days", id = "pseudonym",
                           n.latent = 2, 
                           latentNames = c("symptom_load", "defusion"),
                           n.manifest = 2, 
                           manifestNames = c("A1b", "A3"),
                           LAMBDA = diag(2),
                           DRIFT = matrix(0, nrow = 2, ncol = 2))
head(model_defus_d_0)

# Model fitting
fit_defus_ml_ri_0 <- ctStanFit(datalong = dat_aggregated_c, 
                              ctstanmodel = model_defus_d_0, 
                              priors = FALSE, optimize = TRUE) 

# Summary
summary(fit_defus_ml_ri_0)

# model comparison
ctChisqTest(fit_defus_ml_d_ri, fit_defus_ml_ri_0)

```

## Brain dump

```{r, eval=FALSE}
pred <- c("A1a", "A1b", "A3", "A9", "A4", "A8", "A6", "A2", "A5", "A7")
# Model specification
model_all <- ctModel(type = "stanct", time = "time_hours", id = "pseudonym",
                   n.latent = length(pred)-1, latentNames = c("A1", "A3", "A9", "A4", "A8", "A6", "A2", "A5", "A7"),
                   n.manifest = length(pred), manifestNames = c("A1a", "A1b", "A3", "A9", "A4", "A8", "A6", "A2", "A5", "A7"),
                   LAMBDA = matrix(c(1,1, rep(c(rep(0, length(pred)), 1), length(pred)-2)), nrow = length(pred)))
model_all$pars$indvarying <- FALSE

# Model fitting
fit_all <- ctStanFit(datalong = dat_aggregated_c, ctstanmodel = model_all, priors = TRUE, optimize = TRUE) 
# priors: True -> HMC, false -> max likelihood

# Summary
summary(fit_all)

```


```{r}
dat_aggregated_c %>%
  group_by(pseudonym) %>%
  arrange(time_hours) %>%
  mutate(interval = time_hours - lag(time_hours)) %>%
  summarise(max_interval = max(interval, na.rm = TRUE))

# nlcontrol = list(maxtimestep = 15)
```

